---
output: github_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```

# fragquaxi <img src='man/figures/logo.svg' align="right" height="139" />


## Overview
  
`fragquaxi` allows you to obtain **fr**actional **a**bundances of **g**lycoforms (and proteoforms in general) from mass spectrometric (MS) data by **qua**ntification *via* **X**IC (extracted ion current) **i**ntegration.


## Installation

```{r, eval = FALSE}
devtools::install_git(
  "https://gitlab-ci-token:5pxmPK8746GxLmYuzXB6@git.sbg.ac.at/b1044569/fragquaxi.git"
)
```


## Usage

Load mass spectrometric data.

```{r load_ms_data, message=FALSE}
library(fragquaxi)
library(tibble)
library(dplyr)
library(tidyr)
library(ggplot2)

ms_data <- mzR::openMSfile(
  system.file("extdata", "mzml", "mab1.mzML", package = "fragquaxi")
)
```

Define proteoforms.

```{r define_proteoforms}
proteoforms <- tribble(
  ~name,       ~Hex, ~HexNAc, ~Fuc,
  "G0F/G0",       6,       8,    1,
  "G0F/G0F",      6,       8,    2,
  "G0F/G1F",      7,       8,    2,
  "G1F/G1F",      8,       8,    2,
  "G1F/G2F",      9,       8,    2,
  "G2F/G2F",     10,       8,    2,
)
```

Calculate mass-to-charge ratios of proteoform ions in charge states 33+ to 40+

```{r make_ions}
pfm_ions <-
  proteoforms %>%
  calculate_proteoform_masses("C6464 H9950 N1706 O2014 S44") %>% 
  ionize(charge_states = 33L:40L)

pfm_ions
```

Plot these ions (here, only charge state 36+ of scans 126 to 136).

```{r plot_ions, warning=FALSE}
plot_ions(
  ms_data,
  ions = pfm_ions,
  scans = 126:136,
  xlim = c(4100, 4150)
)
```

Quantify these ions via XIC integration.

```{r quantify_ions}
abundances <- quantify_ions(
  ms_data,
  pfm_ions,
  rt_min = 300,
  rt_max = 350
)

abundances
```

Plot abundances.

```{r plot_abundances}
abundances %>%
  as_tibble() %>% 
  unnest(abundance_data) %>% 
  group_by(name) %>%
  summarise(abundance = sum(abundance)) %>% 
  mutate(frac_ab = abundance / sum(abundance) * 100) %>% 
  ggplot(aes(name, frac_ab)) +
  geom_col() +
  xlab("") +
  ylab("fractional abundance (%)")
```
